

---
title: Week 5 - Inference with One Regressor
sitetitle: Week 5
summary: "EC200 - Week 5 Inference (CH5)"

date: "2022-02-08"
start_date: "2022-02-14"
end_date: "2022-02-18"

  
  # Academic page type (do not modify).
type: docs
menu:
  content:
    parent: Course content
    weight: 5
    
output:
  blogdown::html_page:
    toc: true

pdf: /slides/ch5-slides.pdf
thumb: /slides/ch5-slides.png



---

## Overview

As we carry on into the wonderful world of statistical inference, expect flashbacks of our statistics review. That's what the review was for! It's all coming together  :relieved:.

## Reading Guide

**Chapter 5: Hypothesis Tests and Confidence Intervals**

**SW 5.1**	Testing Hypotheses About One of the Regression Coefficients

This is very important stuff. Again, don’t worry about calculating the variance of a beta coefficient by hand. However, note how similar this is to our hypothesis testing in Chapter 3!

**SW 5.2**	Confidence intervals for a regression coefficient

**SW 5.3**	Regression when X is a binary variable 

We are going to use this ALL THE TIME. 

**SW 5.4**	Heteroskedasticity and homoskedasticity 


```{r echo=FALSE}
blogdown::shortcode('tweet', '1316528345224343552')
```


Drop "heteroskedasticity" into any conversation and you're sure to delight. Just one more benefit of EC200. 
The examples and implications of heteroskedasticity are important. 

And, now you'll know to add `, robust` to all your Stata regressions. (*Quick, someone tell Nate Silver!*)

```{r echo=FALSE}
blogdown::shortcode('tweet', '1273493646969647105')
```


```{r echo=FALSE}
blogdown::shortcode('tweet', '1274901353563934726')
```



**SW 5.5**	Theoretical foundations of OLS

Know the Gauss-Markov theorem and related assumptions. Skip "regression estimators other than OLS." The appendix contains a proof of the Gauss-Markov theorm, but we will not cover that.

~~**SW 5.6**~~ Skip this section!

**SW 5.7**	Conclusion

For the good times.

## Slides

`r blogdown::shortcode("slides")`

## Videos
### [Video: Inference with One-Variable Linear Regression](https://youtu.be/DZo5m5q3bmA)
`r htmltools::HTML("{{< youtube DZo5m5q3bmA >}}")`

## Other resources 

 - [EGAP: 10 things to know about hypothesis testing](https://egap.org/resource/10-things-to-know-about-hypothesis-testing/)
 - [EGAP: 10 things to know about statistical power](https://egap.org/resource/10-things-to-know-about-statistical-power/) - gets a bit deeper than we go, but accessible and handy!
 
