

---
title: Week 14 - Instrumental Variables
sitetitle: Week 14
summary: "EC200 - Week 14 Instrumental Variables (CH12)"

date: "2022-04-18"
start_date: "2022-04-25"
end_date: "2022-04-29"

  
  # Academic page type (do not modify).
type: docs
menu:
  content:
    parent: Course content
    weight: 14
    
output:
  blogdown::html_page:
    toc: true

pdf: /slides/ch12-slides.pdf
thumb: /slides/ch12-slides.png



---

## Overview

Let's wrap up with a final hurrah - one more empirical tool to help us identify *causal impacts*.  

## Reading Guide

#### **Chapter 12: Instrumental Variables Regression**

#### **SW 12.1** The IV Estimator with a Single Regressor and a Single Instrument

Make sure you understand the difference between **endogeneous variables** (correlated with error term $u$) and **exogenous variables** (not correlated). The textbook covers two conditions for a valid instrument: **relevance** ($corr(Z_i,X_i) \neq 0$) and **exogeneity** ($corr(Z_i,u_i) = 0$). To make the assumptions of IV explicit, I prefer to include an **exclusion** assumption: that the instrument $Z$ affects $Y$ only through $X$. 
If you come away from this section with the right vocabulary and a good intuition of what an instrument "does," then you're in good shape!

I find the derivation (in the text and video) helpful, but it's not something I'd test you on.



```{r echo=FALSE}
blogdown::shortcode('tweet', '1298665911101136901')
```


#### **SW 12.2** The General IV Regression Model

Now we get a bit technical. The key important things here are the understanding of two-stage least squares - what it is and how it works, plus a good understanding of those assumptions. 

#### **SW 12.3** Checking Instrument Validity 

Make sure you know what a **weak instrument** is, why it's a problem, and how we identify it. Make sure you understand the logic behind whether an instrument violates it's exogeneity assumption. 



```{r echo=FALSE}
blogdown::shortcode('tweet', '1466634426067075077')
```

```{r echo=FALSE}
blogdown::shortcode('tweet', '1428455696538882048')
```


**Skip the test of overidentifying restrictions**


#### **SW 12.4-12.6** The Other Stuff

Good content and examples, but nothing new. 

## Other resources 

Getting your head around IV can be tricky! Here are a few resources if you want to get another take: 

- [Rebecca Barter: Understanding nstrumental variables](http://www.rebeccabarter.com/blog/2018-05-23-instrumental_variables/): Helpful diagrams!
- [Pierre-Louis Vézina:Instrumental Variables - an intuitive guide](http://pierrelouisvezina.weebly.com/uploads/2/3/4/2/2342194/ivs.pdf): Overview plus a few more examples of IVs in the literature 
- Nick Huntington-Klein has the best econometrics resources :fire: :fire:. 
  - Check out his lecture slides on [instrumental variables](https://nickch-k.github.io/EconometricsSlides/Week_08/Week_08_Instrumental_Variables.html#1), an accompanying video lecture [here](https://www.youtube.com/watch?v=6mFNeWHPMYs). 
  - Also, some [fancy animations](https://nickchk.com/causalgraphs.html#iv) to show what IV actually does w/ data

## Slides

`r blogdown::shortcode("slides")`

##  [Video: Introduction to IV](https://youtu.be/MeHHe303JuI )

`r htmltools::HTML("{{< youtube MeHHe303JuI >}}")`


